---
title: "Hot housing market index"
author: "MaryJo Webster"
date:  "Last updated: `r Sys.Date()`"
output:
  html_document: 
    toc: true
    to_depth: 1
    toc_float: true
    theme: cerulean
    
---

```{r setup, include=FALSE}

#testing

#see the load data section for what data needs to be prepped before running this


knitr::opts_chunk$set(echo = FALSE)

#install.packages("DT")
library(readr) #importing csv files
#library(plyr) #needed for stacked bar chart label positions
library(dplyr) #general analysis 
library(ggplot2) #making charts
library(lubridate) #date functions

library(tidyr)
library(janitor) #use this for doing crosstabs

library(scales) #needed for stacked bar chart axis labels
library(knitr) #needed for making tables in markdown page
#library(car)
library(aws.s3) #needed for uploading to amazon S3 server
library(rmarkdown)
library(DT) #needed for making  searchable sortable data tble
library(jsonlite)
library(formattable)
```



 
```{r, warning=FALSE, message=FALSE}
index_all <-  read_csv('./output/index_main_2021.csv')

#pctstarter <-  read_csv('./output/index_starter_2021.csv')

index_moveup <-  read_csv('./output/index_moveup_2021.csv')


index_starter <-  read_csv('./output/index_starter_2021.csv')








```



# Hot housing Index-- full market
Index_rank=This year's ranking on the index<br>
polp= Average percent of seller's original price<br>
ppsf_chg= Pct change in price per square foot median<br>
closed_chg= Pct change in number of closed sales<br>
dom_diff = Change in average days on market compared to year before<br>
starter_pct = Percent of closed sales that were starter homes

```{r}
index_all <-  left_join(index_all, pctstarter, by=c("place"="place"))


index_all_formatted <-  index_all %>% mutate(starter_pct = percent(pctstarter, digits=0), polp_pct = percent(polp, digits=1), ppsf_pctchange_pct = percent(ppsf_pctchange, digits=1), closed_pctchange_pct = percent(closed_pctchange, digits=1), ppsfavg_c = currency(ppsfavg, digits=0), ppsf_c = currency(ppsf, digits=0))


format_table(index_all_formatted  %>% select(place,index_rank,  closed=closed_sales, polp=polp_pct, ppsf_chg=ppsf_pctchange_pct, closed_chg=closed_pctchange_pct, dom_diff, starter_pct) %>%
  arrange(index_rank))


```








# Price per square foot change
ppsfavg = Average ppsf four years prior<br>
ppsf = Most recent year median price per square foot<br>
```{r}
format_table(index_all_formatted  %>% filter(index_rank!="NA") %>% select(place,index_rank, ppsfavg_c ,ppsf_c, ppsf_pctchange_pct) %>%
  arrange(desc(ppsf_pctchange_pct)))


```

# Change in closed sales
closedavg = average number of closed sales over prevous four years<br>
closed_sales = number of closed sales in most recent year
```{r}
format_table(index_all_formatted  %>% filter(index_rank!="NA") %>% select(place,index_rank, closedavg, closed_sales, closed_chg= closed_pctchange_pct) %>%
  arrange(desc(closed_chg)))


```

# Change in days on market
dom_lastyr = days on market in previous year<br>
dom = days on market in most recent year<br>
dom_diff = difference (in days) between the two years
```{r}
datatable(index_all_formatted  %>% filter(index_rank!="NA") %>% select(place,index_rank, closed_sales, dom_lastyr, dom, dom_diff) %>%
  arrange((dom_diff)))
```



# Starter home market index
This analysis only represents homes less than $300,000
```{r}


index_starter_formatted <-  index_starter %>% mutate(polp_pct = percent(polp, digits=1), ppsf_pctchange_pct = percent(ppsf_pctchange, digits=1), closed_pctchange_pct = percent(closed_pctchange, digits=1), ppsfavg_c = currency(ppsfavg, digits=0), ppsf_c = currency(ppsf, digits=0))

format_table(index_starter_formatted  %>% select(place,index_score,  closed_sales, polp_pct, ppsf_pctchange_pct, closed_pctchange_pct, dom, dom_diff) %>%
  arrange(desc(index_score)))

```

# Move up market index
This analysis represents homes 300k to 500k (hoping to change it to 800k)
```{r}


index_moveup_formatted <-  index_moveup %>% mutate(polp_pct = percent(polp, digits=1), ppsf_pctchange_pct = percent(ppsf_pctchange, digits=1), closed_pctchange_pct = percent(closed_pctchange, digits=1), ppsfavg_c = currency(ppsfavg, digits=0), ppsf_c = currency(ppsf, digits=0))

format_table(index_moveup_formatted  %>% select(place,index_score,  closed_sales, polp_pct, ppsf_pctchange_pct, closed_pctchange_pct, dom, dom_diff) %>%
  arrange(desc(index_score)))


```


# Overall inventory- biggest declines
```{r}
inventory_formatted <-  inventory %>% mutate(inv_chg = percent(pctchange, digits=1))

format_table(inventory_formatted %>%  select(place=name_in_realtors_data, location, x2017, x2018, x2019, inv_chg) %>% arrange((inv_chg)))
```






```{r loadtoaws, echo=FALSE, eval=FALSE, results="hide", message=FALSE, warning=FALSE}

#Load to this URL: http://strib-data-internal.s3-us-west-1.amazonaws.com/projects/HotHousingIndex/HotIndex.html
#this section will not run automatically when you run the whole page



Sys.setenv("AWS_ACCESS_KEY_ID" =  rstudioapi::askForPassword("AWS_ACCESS_KEY_ID"),

           "AWS_SECRET_ACCESS_KEY" = rstudioapi::askForPassword("AWS_SECRET_ACCESS_KEY"))
		   
		   get_bucket("strib-data-internal")


put_object(file = "HotIndex.html", object = "projects/HotHousingIndex/HotIndex.html", bucket = "strib-data-internal")

put_object(file = "HotIndex.html", object = "projects/HotHousingIndex/HotIndex.html", bucket = "strib-data-internal", acl=c("public-read"))
```

